### 2018-08-17 submission_1/codeForPaper/newcomb/directSampling/readMeDirectSampling.tex
added code for first draft submission to the repo.

readMe: DirectSampling

Direct Sampling methods for evaluated the restricted posterior [theta, sigma^{2}| T(X)] for T(X) = Huber/Huber Proposal 2, LTS, and LMS estimators

##
speedOfLightAnalysisHuberDirectSampling
description: Posterior evaluation using Huber's estimators. Saves results to a workspace  wsNewcombDirectSamplingHuber2.RData

##
speedOfLightLMSDirectSampling
description: Posterior evaluation using LMS. Saves results to a workspace  wsNewcombDirectSamplingLMS2.RData

##
speedOfLightLTSDirectSampling
description: Posterior evaluation using LTS. Saves results to a workspace  wsNewcombDirectSamplingLMS2trim_2.RData
*Note: trim_2 refers to k=62, the value ultimately used in the paper. This can be changed in the code to use k=34. If done, the ws is saved as wsNewcombDirectSamplingLMS2.RData