### 2018-08-17 submission_1/codeForPaper/newcomb/mcmc/readMeMCMC.tex

ReadMe: newcomb/mcmc

##
locAndScaleHuberAndProposal2 and locAndScaleTukeyAndProposal2

These are source files containing functions used to sample from the restricted posterior using the M-estimators used in the paper. They are sourced by newcombHuberAndProposal2Analysis.R and  newcombTukeyAndProposal2Analysis.R, respectively

*Note: These are specific to the location and scale setting. Functions have since been updated to allow for covariates and provided in linearRegressionHuberAndProposal2.

##
newcombFullNormalModel.R
Fits the full normal theory model. Saves results in workspace WorkSpaceNewcombFullNormalModel2.RData

##
newcombHuberAndProposal2Analysis.R and  newcombTukeyAndProposal2Analysis.R

Uses mcmc to fit the restricted models. Also includes analysis done where the outliers are deleted (not discussed in the paper). Workspaces are saved in WorkSpaceNewcombHuberAndProposal2AnalysisWithDelAnalysis2.RData and WorkSpaceNewcombTukeyAndProposal2AnalysisWithDelAnalysis2.RData.