### 2018-08-17 submission_1/codeForPaper/newcomb/readMeNewcomb.tex

ReadMe: Newcomb

Code to analyze the speed of light data in several ways

##

newcombAnalyzeOutput.R
description: Creates plots used in paper. Must have access to the workspaces created by code in the directSampling and mcmc folders.

##

newcombFixedParameters1.R and newcombFixedParameters2.R
description: defines hyper parameters. Two different versions. Values in newcombFixedParameters2.R were ultimately used in the paper.

##
Folders

##
directSampling: 
Uses 'direct sampling' techniques to sample from the restricted posterior under Huber, LTS, and LMS. See read me file in the folder

##
fittingTdist
uses MCMC and grid techniques to fit the t-distribution model.  See read me file in the folder

##
mcmc
uses MCMC techniques to sample from the full posterior and restricted posterior under Huber/huber prop2 and Tukey/Huber prop2 estimators. See read me file in the folder