### 2018-08-17 submission_1/codeForPaper/readMeCodeForPaper.tex

ReadMe: codeForPaper

##
linearRegressionHuberAndProposal2.R
  description: contains functions for the MH algorithm for sampling from the full conditional [Y|T(Y)=T(y_obs), theta] for the Bayesian linear model. This is perhaps the most important piece of code.

NOTE: the original file was named as such because it was originally written to only handle Huber's estimators. It has morphed into slightly more than that since.

##
bayesLM.R
	description: contains functions for MCMC algorithms for various linear models
	bayesLm: standard normal theory regression. See file for full description of model. 
	restrictedBayesLm: linear model conditioning on summary statistics.
	bayesTdistLm: linear model using Student's t-distribution. 
	bayesTdistLm2: second version of MCMC for the t model. Same stationary distribution just different sampling method.

##Folders##

##	
nwLinearReg
	See readMeLinearReg in folder
##	
nwHierLinearReg
	See readMeHierLinearReg in folder

## 
newcomb
	See readMeNewcomb in folder

##
simulation
  See readMeNewcomb in folder