### 2018-08-17 submission_1/codeForPaper/simulation/readMeSimulation.tex

readMe: Simulation

##
simulation.R: code for simulation described in paper. Saves results in simulationWorkspace.RData or simulationWorkSpace_concTau2Near1.RData. The latter uses a prior that concentrates around the true value for tau^{2} of 1. These are not provided as they are large workspaces.

##
simulationSourceCode.R: code sourced by simulation.R. This defines functions necessary to fit each model. Also defines a function to simulation data (fn.gen.data)

##
postProcessingSimulation.R
post-processes simulation results and computes evaluation metrics. Saves results in simPostProcessed.RData or simPostProcessed_concTau2Near1.RData.

##
summarizeSimulations.R
summarizes the evaluation metrics computed in postProcessingSimulation.R and creates some plots

##
testSimulationSourceCode.R
some code to informally check that the functions in simulationSourceCode.R are giving reasonable results.